COG-ED Manuscript

### 2021-10-22 06_Paper/COG-ED/Manuscript.Rmd
fit mlm model with lmerTest

  knitr::opts_chunk$set(echo = TRUE)
  library(bibtex)
  library(here)         # to set a directory without defined paths, "here" package
  library(tidyverse)    # plotting and managing data, "tidyverse" package
  library(bayestestR)   # computing area under the curve, "bayestestR" package
  library(papaja)       # APA conform output, "papaja" package
  library(nlme)         # fitting nonlinear mixed-effects model
  library(afex)         # compute repeated measures ANOVA, "afex" package
  library(emmeans)      # for custom contrasts in ANOVAs, "emmeans" package

@@

## Analysis
We aimed to conduct all analysis as described in Westbrook et al. [-@Westbrook2013], but the level of detail was not always sufficient, so there might be deviations in terms of data cleaning and degrees of freedom in the statistical procedures.
The performance measure *d'* was computed as the difference of the *z*-transformed hit rate and the *z*-transformed false alarm rate [@Macmillan1990].
Reaction time (RT) data was trimmed by excluding all trials with responses faster than 200 ms, as the relevant cognitive processes could not have been completed before [@Ashby1980; @Whelan2008; @Berger2021].
Post-error trials were also excluded, because RT on those trials is longer due to more cautious behavior [@Dutilh2012; @Houtman2012].
Declining performance was investigated by calculating a repeated measures analysis of variance (rmANOVA) with three paired contrasts comparing *d'* between two levels of 2-, 3-, and 4-back at a time.
Another rmANOVA with three paired contrasts was computed to compare the mean RT between two levels of 2-, 3-, and 4-back at a time.
To investigate changes in NASA-TLX ratings, six rmANOVAs were computed, one for each NASA-TLX subscale, and each with six paired contrasts comparing the ratings between two levels of 1-, 2-, 3-, and 4-back at a time.
The SVs were calculated by adding or subtracting the final iteration value of 0.015625 from the monetary value of the flexible level for each ED round, depending on the participant's choice in the last iteration.
Then, the final monetary values were divided by 2€, and the SV of each level per participant was computed by averaging all final monetary values of each level, regardless of whether it was fixed or flexible.
A rmANOVA with six paired contrasts was computed, comparing the SVs between two levels of 1-, 2-, 3-, and 4-back at a time.
Tukey method was used for the paired contrasts of each rmANOVA, including *p*-value adjustment.
To determine the influence of task performance in the association of SVs and n-back level, we set up a multi-level model.
**MLM explanation goes here**
The association of ED and NFC was examined using three separate regressions.
One regression used the AUC of each participant's SVs to predict their NFC score.
The second regression additionally included the mean of the NASA-TLX subscales' AUCs of each participant as a predictor.
And the last regression **insert method**.

@@

  # define variables -------------------------------------------------------------
  # dependent variable:     SV                    numeric                 data_SV
  # indepenendent variable: n-back level [1-4]    numeric                       
  # controlled regressors:  d'                    numeric                 dprime
  #                         mRT                   numeric                 rt

  # get all relevant variables ---------------------------------------------------

@@

  # check and specify correct data types -----------------------------------------
  str(h2b_data)

  # center predictors ------------------------------------------------------------
  # center level 1 predictors
  # two ways: centering at the grand mean (CGM) vs. centering within cluster (CWC) = group mean centering
  # level 1 predictor (SV) is of substantive interest
  # > CWC

@@

  # no level 2 predictors

  # change global coding to orthogonal sum-to-zero contrast by -------------------
  options(contrasts = c("contr.sum", "contr.poly"))

  # null model -------------------------------------------------------------------
  # m0_h2b <- nlme(model = sv ~ 1,
  #                data = h2b_data,
  #                random = ~ 1 | subject, 
  #                method = "REML")

  # full model -------------------------------------------------------------------
  # random intercept vs. random slope ???
  # "subject-specific intercepts" speaks for random-slopes-model, as within this model 
  # the slope is allowed to vary between subjects

  # m1_h2b <- nlme(sv ~ d.cwc + meanrt.cwc + level.cwc + (level.cwc| subject),
  #                data = h2b_data,
  #                fixed = ...,
  #                random = ...,
  #                control = ...,
  #                method = "REML")

  # initVals <- getInitial(sv ~ d.cwc + meanrt.cwc + level.cwc, data = h2b_data)

  # get initial values for mlm start
  fm.lm <- lm(sv ~ d.cwc + meanrt.cwc + level.cwc, h2b_data)
  st <- coef(fm.lm)

  m1_h2b <- nlme(sv ~ d.cwc + meanrt.cwc + level.cwc,
                 data = h2b_data,
                 #fixed = d.cwc + meanrt.cwc + level.cwc ~ 1,
                 fixed = list(d.cwc ~ 1, meanrt ~ 1, level.cwc ~ 1),
                 random = level.cwc ~ 1 | subject,
                 start = list(fixed = st[2:4]),
                 method = "REML")

  # change global coding to default ----------------------------------------------
  options(contrasts = c("contr.treatment", "contr.poly"))

>>>>

  knitr::opts_chunk$set(echo = TRUE)
  library(bibtex)
  library(here)         # to set a directory without defined paths, "here" package
  library(tidyverse)    # plotting and managing data, "tidyverse" package
  library(bayestestR)   # computing area under the curve, "bayestestR" package
  library(papaja)       # APA conform output, "papaja" package
  library(lmerTest)     # for multilevel modeling
  library(snow)         # for faster processing
  library(afex)         # compute repeated measures ANOVA, "afex" package
  library(emmeans)      # for custom contrasts in ANOVAs, "emmeans" package

@@

## Analysis
We aimed to conduct all analysis as described in Westbrook et al. [-@Westbrook2013], but the level of detail was not always sufficient, so there might be deviations in terms of data cleaning and degrees of freedom in the statistical procedures.
The performance measure *d'* was computed as the difference of the *z*-transformed hit rate and the *z*-transformed false alarm rate [@Macmillan1990].
Reaction time (RT) data was trimmed by excluding all trials with responses faster than 200 ms, as the relevant cognitive processes could not have been completed before [@Ashby1980; @Whelan2008; @Berger2021].
Error- and post-error trials were excluded, because RT on the latter trials is longer due to more cautious behavior [@Dutilh2012; @Houtman2012].
#TODO: Tolle Artikel, die schaue ich mir auch noch mal genauer an! Hast du die beiden exclusions schon im script umgesetzt? Falls ja, in welcher Zeile, ich kann es nicht finden...
Declining performance was investigated by calculating a repeated measures analysis of variance (rmANOVA) with three paired contrasts comparing *d'* between two levels of 2-, 3-, and 4-back at a time.
Another rmANOVA with three paired contrasts was computed to compare the mean RT between two levels of 2-, 3-, and 4-back at a time.
To investigate changes in NASA-TLX ratings, six rmANOVAs were computed, one for each NASA-TLX subscale, and each with six paired contrasts comparing the ratings between two levels of 1-, 2-, 3-, and 4-back at a time.
The SVs were calculated by adding or subtracting the final iteration value of 0.015625 from the monetary value of the flexible level for each ED round, depending on the participant's choice in the last iteration.
Then, the final monetary values were divided by 2€, and the SV of each level per participant was computed by averaging all final monetary values of each level, regardless of whether it was fixed or flexible.
A rmANOVA with six paired contrasts was computed, comparing the SVs between two levels of 1-, 2-, 3-, and 4-back at a time.
Tukey method was used for the paired contrasts of each rmANOVA, including *p*-value adjustment.
To determine the influence of task performance in the association of SVs and n-back level, we set up a multi-level model.
**MLM explanation goes here**
The association of ED and NFC was examined using three separate regressions.
One regression used the AUC of each participant's SVs to predict their NFC score.
The second regression additionally included the mean of the NASA-TLX subscales' AUCs of each participant as a predictor.
And the last regression **insert method**.

@@

  # detach afex in order to user lmerTest
  detach("package:afex", unload = TRUE)

  # get all relevant variables -------------------------------------------------

@@

  # center predictors ----------------------------------------------------------

  ## center level 1 predictors
  # two ways: centering at the grand mean (CGM) vs. centering within cluster (CWC) = group mean centering
  # level 1 predictor (SV) is of substantive interest
  # > CWC

@@

  ## no level 2 predictors

  # null model -----------------------------------------------------------------
  m0_h2b <- lmer(sv ~ 1 + (1|subject), 
                 data = h2b_data)

  # intraclass correlation (ICC)
  # % of variance can be explained by differences between persons
  var_m0_h2b <- as.data.frame(VarCorr(m0_h2b))
  icc_h2b <- var_m0_h2b$vcov[1] / (var_m0_h2b$vcov[1] + var_m0_h2b$vcov[2])

  # random slopes model --------------------------------------------------------

  ## maximal model ----

  m1_h2b <- lmer(sv ~ level.cwc + d.cwc + meanrt.cwc + (level.cwc|subject), 
                 data = h2b_data,
                 REML = T) # parameters estimated by maximizing restricted log-likelihood

  # (1) if a warning message is produced, check random effects
    summary(m1_h2b)$varcor
    # (11) if all SDs are different from 0 and correlations look good (unequal to zero)
      # this message should be a false positive and this model can be reported
    # (12) if (11) is not true, check all available optimizers
      n_cores <- parallel::detectCores()
      opt_cl <- makeCluster(n_cores, type = "SOCK")
      clusterEvalQ(opt_cl, library("lme4"))
      clusterExport(opt_cl, "h2b_data")
      m1_h2b_all <- allFit(m1_h2b, parallel = "snow", ncpus = n_cores, cl = opt_cl) 
      stopCluster(opt_cl)
      summary(m1_h2b_all)$msgs
      # (121) if an optimizer returns NULL, use this optimizer and recalculate model to get the final model

  # (2) if (121) is not true OR a singular fit message occurs, check for potential outliers and recalculate
    potential_outliers_m1_h2b <- subset(h2b_data, abs(scale(resid(m1_h2b))) > 3)    
    percentage_outliers <- 100/nrow(h2b_data) * nrow(potential_outliers_m1_h2b)
    # get new subset excluding outliers
    h2b_data_norm <- droplevels(subset(h2b_data, abs(scale(resid(m1_h2b))) < 3))
    # start with (1) again

  # (3) if (1) and (2) do not help, start reducing the model stepwise
